---
title: "Statistical Inference Course Project Part 1"
author: "domesc"
date: "June 16, 2015"
output: pdf_document
---


### Assignment
In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials.

### Analysis
First we need to set up the variables and simulations to be used for the assignment
```{r}
library(ggplot2)
set.seed(42)
lambda <- 0.2
nosim <- 1000
n <- 40

sim <- data.frame(mean=numeric(nosim))
for (i in 1 : nosim) {
  sim[i,1] = mean(rexp(n, lambda))
}
```


### 1. Show the sample mean and compare it to the theoretical mean of the distribution.

We compute the theoretical mean through the formula and we use the R function "mean" to compute the sample mean
of our simulations vector:
```{r}
theo_mean <- 1/lambda
sample_mean <- mean(sim$mean)

theo_mean
sample_mean
```

We can see also with a plot that the two means are quite close:
```{r}
plot <- ggplot(data = sim, aes(x=sim$mean))+
  ggtitle("Distribution of sample means") +
  xlab("Sample Means from 1000 simulations") +
  ylab("Frequency") +
  geom_histogram(aes(y=..density..), color = 'black', fill = 'white', binwidth = 0.25) 

plot +
  geom_vline(aes(xintercept=theo_mean, color='Theoretical'), show_guide=T) + 
  geom_vline(aes(xintercept=sample_mean, color='Sample'), show_guide=T) +
  scale_color_manual("Mean",
                     values=c("Theoretical"="blue","Sample"="red"))
```

### 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

We do the same for the variance:
```{r}
theo_var <- (1/(lambda^2))/n
sample_var <- var(sim$mean)

theo_var
sample_var
```

### 3. Show that the distribution is approximately normal.

We plot the histogram of our distribution and compare it with the normal distribution associated to 
the theoretical mean:
```{r}
plot +
  geom_line(aes(x = sim$mean,y = ..density.., colour = 'Simulation'), stat = 'density') +  
  stat_function(
    fun = dnorm, args = list(mean = theo_mean, sd = sqrt(theo_var)), 
    aes(colour = 'Theoretical')
    ) +
  theme(legend.title=element_blank())
```

The plot shows well that our distribution is close to a normal distribution.
