---
title: "Statistical Inference Course Project Part 2"
author: "domesc"
date: "June 16, 2015"
output: pdf_document
---

### Assignment
Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package.

### 1.Load the ToothGrowth data and perform some basic exploratory data analyses
We load first the data and we look at the variable types through str function:
```{r}
data("ToothGrowth")
str(ToothGrowth)
```
We have 2 numerical columns "len" and "dose" and a factor variable "supp". To have more information on the
data set:
```{r}
head(ToothGrowth)
help("ToothGrowth")
summary(ToothGrowth)
```

### 2.Provide a basic summary of the data
```{r}
library(ggplot2)
plot <- ggplot(data = ToothGrowth, aes(x=ToothGrowth$dose, y = ToothGrowth$len))+
  ggtitle("Scatterplots by supplement type") +
  xlab("Dose in milligrams") +
  ylab("Tooth length") +
  geom_point(color = 'black', fill = 'white') +
  facet_wrap(~supp)
plot
```

### 3.Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

First of all we use the t.test to check if the difference of Vitamin C supplement type influence the tooth growth. We do this by looking at the p value:
```{r}
t.test(ToothGrowth$len[ToothGrowth$supp=="OJ"], 
       ToothGrowth$len[ToothGrowth$supp=="VC"], 
       paired = FALSE, 
       var.equal = FALSE)
```
In this case it is not < 0.05, so we cannot reject the null hypothesis and we can conclude that there is no evidence for a correlation between supplement type and tooth growth.

We now check if the dose can influence tooth growth always by looking at the p value:
```{r}
t.test(ToothGrowth$len[ToothGrowth$dose==2], 
       ToothGrowth$len[ToothGrowth$dose==1], 
       paired = FALSE, 
       var.equal = TRUE)
```
this time since the p value is <0.05 we have enough evidence to reject the null hypothesis. This means that there is a correlation between a dose change from 1 to 2 mg and the tooth growth.

```{r}
t.test(ToothGrowth$len[ToothGrowth$dose==1], 
       ToothGrowth$len[ToothGrowth$dose==0.5], 
       paired = FALSE, 
       var.equal = TRUE)
```
The confidence interval does not contain zero. There is no need to compare the dosage of 0.5 and 2 mg.


### 4.State your conclusions and the assumptions needed for your conclusions.
We can conclude saying that there is no correlation between supplement type and tooth growth. On the other hand while increasing the dosage of vitamin C we can see a growth in the guinea pig's teeth.
